\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{enumitem}

\begin{document}

\begin{center}
{\Large \textbf{In-Class Exercise (Canvas Quiz) --- 10 minutes}}\\[6pt]
{\large \textbf{Lecture 17: Best Practices}}
\end{center}

\vspace{6pt}
\noindent
\textbf{Canvas instruction:} Canvas will show only keywords. Full questions are shown on the projector.

\vspace{8pt}
\noindent
This exercise is based directly on the end-of-lecture questions in Lecture 17 (Best Practices): selecting losses and debugging exploding gradients.

\vspace{10pt}
\noindent
\textbf{Tasks (answer in Canvas):}
\begin{enumerate}[label=(\arabic*), itemsep=8pt]

\item \textbf{Loss for balanced 10-class classification.}\\
Choose the best loss function:
\begin{itemize}
    \item A. Mean Squared Error (MSE)
    \item B. Cross-Entropy
    \item C. L1 Loss
    \item D. Weighted Cross-Entropy
\end{itemize}

\item \textbf{Loss for imbalanced tumor segmentation.}\\
Choose the best loss function:
\begin{itemize}
    \item A. Cross-Entropy
    \item B. Weighted Cross-Entropy
    \item C. MSE
    \item D. Focal Loss
\end{itemize}

\item \textbf{Fix exploding gradients via activation histogram.}\\
What is the best fix?
\begin{itemize}
    \item A. Increase learning rate
    \item B. Decrease learning rate
    \item C. Use a more aggressive activation like ReLU
\end{itemize}

\end{enumerate}

\vspace{6pt}
\noindent
\textbf{Time limit:} 10 minutes total.

\end{document}
